Using best practices for forecasting biodiversity changes in breeding birds

### 2017-07-11 manuscript/draft.Rmd

Forecasting is an emerging imperative in ecology and as such the field needs to
develop and follow good practices for conducting and evaluating ecological
forecasts. We have laid out a number of these practices (Box 1) and attempted to
implement them in a single study that builds and evaluates forecasts of
biodiversity in the form of species richness. The results of this effort are
both promising and humbling. On an absolute basis, when comparing forecasts
across sites, many different approaches to forecasting produce reasonable
forecasts. If a site is predicted to have a high number of species in the
future, relative to other sites, it generally does. However, the forecasts
perform relatively poorly at predicting variation in richness over time for any
given site. If a site is predicted to have a higher richness next year than it
did last year there is little confidence that this will be the case. As a
result, the baseline models, like the long-term average and the "naive" model,
along with site-level time-series models, provided some of the best forecasts
for future biodiversity.

The most commonly used method to forecast future biodiversity, stacked SDMS
[[cite?]], provided the worst overall forecasts of all methods evaluated. In
general, SDM-based forecasts are rarely evaluated, particularly in terms of
repeated observations at the same site, and especially with long time horizons.
Since site-level richness is relatively stable, the model evaluations that show
high accuracy in predicting differences among sites may not indicate much about
the model's ability to predict changes over time. This issue has also recently
been recently highlighted as an issue for using SDMs to predict the future
locations of individual species [plos one 2012 paper]. This result is
particularly sobering because this approach forms the foundation for major
claims regarding the predicted loss of biodiversity to climate change [2004
science/nature papers]. The poor performance of SDM-based forecasts highlights
the crucial importance of comparing multiple modeling approaches when conducting
ecological forecasts, and in particular the value of comparing results to simple
baselines to avoid over-interpreting the accuracy of ecological forecasts.
Disciplines that have more mature forecasting cultures often do this by
reporting "forecast skill", which is the improvement in the forecast relative to
a simple baseline. We recommend adopting this approach in future ecological
forecasting research. It also highlights the need for using hindcasting and
time-series data for assessing the effectiveness of forecasting methods, since
this allows comparisons to baselines and assessment of the predicted dynamics at
each site.

The most common alternative to summed SDMs is modeling richness directly using
environmental variables and using the resulting models to make forecasts []. Two
comparisons of the the SDM and direct richness modeling approaches reported that
the methods yielded equivalent results for forecasting diversity. While our
results generally support rough equivalence between stacked SDMs and direct
richness modeling for point estimates, they also show that modeling richness
directly results in much better estimation of the uncertainty of the forecast
and therefore is a better overall approach to forecasting richness. A similar
result is seen when comparing joint species distribution models (JSDMs) to
stacked single species distribution models. For point estimates the joint
distribution models are roughly equivalent to stacked SDMs, but the JSDMs
provide much better estimates of uncertainty. In fact, JSDMs and modeling
richness directly provide some of the best estimates of uncertainty across all
modeling approaches. This highlights the importance evaluating models based on
their uncertainty as well as their mean estimates. Not doing so is particularly
problematic in this case because the summed species distribution models are
overly confident meaning that using them would lead to believing that richness
would be restricted to a much narrower range than would actually occur.

For error metrics that evaluate both models' point estimates and their
uncertainty, the joint SDM and the nonlinear richness model began to outperform
the time-series baselines at the longest time-scales assessed. This convergence
at longer time scales highlights the importance of considering how forecasting
method performance changes with the distance into the future being predicted.
The value of this approach, and associated forecast horizons, has been raised
for knowing how far into the future a model can be effectively used for
forecasting [@petchy2015], but our results suggest a broader value to this
approach for considering the potential importance of different models and
processes for making forecasts at different scales. [[Should make sure Petchy et
al. don't make this point]] Our results show that for forecasts 1-3 years in the
future, baselines and time-series based approaches outperform all other methods
using integrative metrics. At these time scales, environmental changes are
relatively small and changes in the biota may lag behind changes in the
environment. However, as the timescale of the forecast increases to a decade the
JSDMs and environmental richness models get closer to the accuracy of the
time-series methods and even surpass some of those methods. As the forecast
distance increases the amount of environmental change is expected to increase
and the system will have more time to respond, potentially leading to shifts
towards improved relative performance of the models incorporating environmental
and ecological information. If this shift with forecast distance in the relative
importance of different processes and models continues, this could lead to JSDMs
and environmental richness models outperforming time-series based approaches at
sufficiently long time scales. These results suggest that assessment of how
forecast performance changes with time lag is not only important for determining
how far in the future to forecast, but also for determining models and processes
are most relevant for making forecasts at different time scales.

It is also possible that models that include species' relationships to their
environments or direct environmental constraints on richness may continue to
produce forecasts that are worse than simply assuming the systems are
static. This would be expected to occur is the systems are in fact not changing
their richness over the relevant time scales and therefore simple models of no
directional change are appropriate. Recent suggestions that local scale richness
in some systems is not changing directionally at multi-decadal scales supports
this possibility [@brown2001; @ernest2001; @dornelas201X; @theplantone]. This
lack of change may be expected even in the presence of substantial changes in
environmental conditions and species composition at a site due to the
replacement of species from the regional pool [@brown2001; @ernest2001]. On
average the Breeding Bird Survey sites used in this study show little change in
richness ([[Add summary stat here]]) [see also @lasorte2007]. This absence of
rapid change is beneficial for the absolute accuracy of forecasts across sites,
because if a past years richness is known it is easy to estimate a future
richness, and explains why stacked SDMs performs relatively well at this task
despite failing to capture meaningful dynamics. However it makes it difficult to
improve forecasts relative to simple baselines, since those baselines are close
to representing what is actually occurring in the system. This suggests that
simple time-series models and baselines should be actively considered for
forecasts of richness and other stable aspects of biodiversity and also suggests
that future efforts to understand and forecast biodiversity should also focus on
composition since it is expected to be more dynamic
[@ernest2001; @dornelas201X].

In addition to consideration of the different process models used for
forecasting it is important to consider the observation models. When working
with any ecological dataset there are imperfections in sampling that have the
potential to influence results. With large scale survey and citizen science
datasets like the Breeding Bird Survey these issues are potentially magnified by
the large number of different observers and major differences in habitat and
species. We included an observation model for one of the two major observation
issues known for the Breeding Bird Survey, differences among
observers. Accounting for differences in observers substantially reduced the
error in point estimates in all models and also improved the coverage of the
confidence intervals. In addition, it resulted in changes in which models
performed best, most notably reducing the performance of the naive model for
point estimates. This suggests that the naive model performed well in part
because it was capable of accommodating rapid shifts in estimated richness
introduced by changes in the observer. These kinds of rapid changes were
difficult for the other time-series models to accommodate and so the ARIMA and
average models improved substantially once this source of observation error was
addressed. This demonstrates that properly modeling observation error can be
important for reducing uncertainty in forecast but can also lead to changes in
the best methods for forecasting. We did not address differences in detection
probability across species and sites since there is no clear way to address this
issue without making strong assumptions about the data, but this would be a
valuable addition to future forecasting models.

Here we have made forecasts using yearly timesteps fifty (?) years into the
future. Many ecological forecasts are temporally aggregated to 5-30 year
timesteps and projected up to a century into the future. These are commonly
beyond the career or even lifespan of the researchers, making reasonable model
assessment impossible. Currently assessments at these large scales can be only
be made with a small number of opportune datasets. For examples hindcasts have
been used to asses models of species richness [@alger2009, distler2015] and
distribution [@rappacciuolo2012, @moran-ordonez2017, @araujo2005,
@eskildsen2013] using data aggregated across similar time scales. More
evaluation studies like these are needed but there is a paucity of adequate long
term data available. Short term forecasts will allow for relatively rapid
validation of models by researchers without needing to wait for enough data to
accumulate. This is not to say they should replace large scale long term
forecasts. The relationships of temporal scaling is currently understudied, but
it is possible they are similar to spatial scale dynamics [@mcgill2010,
@levin1992]. For example drivers of species richness differ at different spatial
scales [@rahbek2001, @hurlbert2003]. Thus short term forecasts should be
expected to add value to end users and inform, not replace, long term larger
scaled forecasts.

Future efforts will also need to begin to address the additional uncertainty
that comes from error in forecasting the environmental conditions themselves. In
this and other hindcasting studies the environmental conditions for the "future"
are known because the data has already been observed. In real forecasts the
environmental conditions themselves have to be forecast and those forecasts have
uncertainty and bias. This will cause ecological forecasts that use
environmental data to be more uncertain. It is important to correctly
incorporate this uncertainty in the predictor variables into forecasting models
[@clark2001; @dietz2017]. Difficulty in forecasting future environmental
conditions at small scales will present continued challenges for models
incorporating these conditions and this may result in a continued advantage to
simple time-series based approaches.

The science of forecasting biodiversity, and ecology more broadly, remains in
its infancy and it is important to consider the general inability the
forecasting methods to improve on simple baselines in that context. When weather
forecasting first started the forecasts were likewise worse than simple
baselines []. One of the things that helped weather forecasts improve was large
numbers of forecasts were made in public, which allowed different approaches to
forecasting to be regularly assessed and improved
[@mcgill2012 blog post; history of weather forecasting book]. This suggests that
it is important for ecologists to start regularly making and evaluating true
ecological forecasts, even if they perform poorly, and to make these forecasts
publicly available for assessment. These forecasts should include both
short-term predictions, which can be assessed quickly, and mid to long-term
forecasts to help assess long time-scale processes and determine how far into
the future we can successfully forecast [@tredenick ???, @dietz2017]. Forecasts
for the next XX years for the models in this paper are openly archived on Dryad
(XXXXXXXXXXX) so that we and others can assess how well they perform and we plan
to evaluate these forecasts and report the results as each new year of BBS data
becomes available. Weather forecasting has continually improved throughout its
history [@bauer2015] in part due to making and evaluating public forecasts
[@mcgill2012] and we hope this will help ecology do the same.

Making successful ecological forecasts will be challenging. Ecological systems
are complex, fundamental theory is less refined than simpler physical/chemical
systems, and we currently lack the kinds of truly "big" data that produce
effective forecasts through machine learning. Despite this we believe that
progress can be made if we build an active forecasting culture in ecology that
builds and assesses forecasts in ways that will allow us to improve the
effectiveness of ecological forecasts most rapidly (Box 1). This includes
expanding the scope of the ecological and environmental data we work with,
paying attention to uncertainty in both model building and forecast evaluation,
and rigorously assessing forecasts using a combination of hindcasting, archived
forecasts, and comparisons to simple baselines.

@@

Typically ecological forecasts use one modeling approach or a small number of
related approaches. By fitting and evaluating multiple modeling approaches we
can learn more rapidly about the best approaches for making predictions for a
given ecological quantity. This includes comparing process based and data-driven
models and comparing the accuracy of forecasts to simple baselines to determine
if the modeled forecasts are more accurate than the naive assumption that the
world is static.

@@

Forecasts describe how systems are expected to change through time. While some
areas of ecological forecasting focus primarily on time-series data, others
primarily focus on using spatial models and space-for time substitutions. Using
ecological and environmental time-series data allows the consideration of actual
dynamics from both a process [@treddenick2016a] and error structure perspective.

@@

Understanding confidence in a forecast is just as important as understanding the
average or expected outcome. Failing to account for uncertainty can result in
overconfidence in highly uncertain outcomes leading to poor decision making and
erosion of confidence in ecological forecasts. Models should explicitly include
sources of uncertainty and how the propagate through the forecast where
possible. Evaluations of forecasts should assess the accuracy of uncertainties
as well as point estimates [@dietz2017].

@@

Many ecological forecasts use data that is readily available and easy to work
with. While ease of use is a reasonable consideration it is also important to
include predictor variables that are expected to relate to the ecological
quantity being forecast and dynamic time-series of predictors instead of
long-term averages. Investing time in identifying and acquiring better predictor
variables may have at least as many benefits as using more sophisticated
modeling techniques.

### 5. Assess how forecast accuracy changes with time-lag

@@

### 6. Include an observation model

Ecological observations are influenced by both the underlying processes and how
the system is sampled. When possible forecasts should model the factors
influencing the observation of the data.

### 7. Validate using hindcasting

To evaluate the expected accuracy and uncertainty of forecasts assess the
performance of these forecasts within existing time-series data.

### 8. Publicly archive forecasts

To allow the future evaluation of the accuracy and uncertainty of forecasts the
forecast values and/or models should be archived so that they can be assessed
after new data is generated [@mcgill2012]. Enough information should be provided
to allow an unambiguous assessment of the forecast performance.

### 9. Make short-term and long-term predictions

In cases where long-term predictions are the primary goal, short-term should
also be made to accommodate the time-scales of planning and management decisions
and to allow the accuracy of the forecasts to be quickly evaluated
[@treddenick2016a].

### 10. [[?]]

Sites' biotic and abiotic environments differ in more ways than the 
few ones you happened to measure. As a result, some sites will 
consistently have more species than a regression model (or sum of 
regression models) would predict, while others willl consistently have 
fewer. Depending on the spatial scale of these unmeasured variables, 
independent site-level random effects could be sufficient for dealing 
with this, or spatially-autocorrelated random effects might be needed.

>>>>

Forecasting is an emerging imperative in ecology and as such the field needs to
develop and follow best practices for conducting and evaluating ecological
forecasts. We have laid out ideas for a number of these practices (Box 1) and
attempted to implement them in a single study that builds and evaluates
forecasts of biodiversity in the form of species richness. The results of this
effort are both promising and humbling. On an absolute basis, when comparing
forecasts across sites, many different approaches to forecasting produce
reasonable forecasts. If a site is predicted to have a high number of species in
the future, relative to other sites, it generally does. However, the forecasts
perform relatively poorly at predicting variation in richness over time at a
location. If a site is predicted to have a higher richness next year than it did
last year there is little confidence that this will be the case. As a result,
the baseline models, like the long-term average, along with site-level
time-series models, provided some of the best forecasts for future biodiversity.

The most commonly used method to forecast future biodiversity, stacked SDMS
[@urban2015], provided the worst overall forecasts of all methods evaluated. In
general, SDM-based forecasts are rarely evaluated, particularly in terms of
repeated observations at the same site. Since site-level richness is relatively
stable, model evaluations that show high accuracy in predicting differences
among sites may not indicate much about a model's ability to predict changes
over time. This issue has also recently been recently highlighted as an issue
for using SDMs to predict the future locations of individual species
[@rapacciuolo2012]. This result is particularly sobering because this approach
forms the main foundation for estimates of the predicted loss of biodiversity to
climate change [@thomas2004; @urban2015]. The poor performance of SDM-based
forecasts highlights the crucial importance of comparing multiple modeling
approaches when conducting ecological forecasts, and in particular the value of
comparing results to simple baselines to avoid over-interpreting the information
present in these forecasts (Box 1).  Disciplines that have more mature forecasting
cultures often do this by reporting "forecast skill", which is the improvement
in the forecast relative to a simple baseline [[weather citation]]. We recommend
following Ye et al. [@ye2015] and adopting this approach in future ecological
forecasting research. It also highlights the need for using hindcasting and
time-series data for assessing the effectiveness of forecasting methods, since
this allows comparisons to baselines and assessment of predicted dynamics within
sites or regions.

One of the main alternatives to summed SDMs is modeling richness directly using
environmental variables and using the resulting models to make forecasts
[@algar2009; @distler2015]. Two comparisons of the the SDM and direct richness
modeling approaches reported that the methods yielded equivalent results for
forecasting diversity. While our results generally support rough equivalence
between stacked SDMs and direct richness modeling for point estimates, they also
show that modeling richness directly results in much better estimation of the
uncertainty of the forecast and therefore is a better overall approach to
forecasting richness. A similar result is seen when comparing joint species
distribution models (JSDMs) to stacked single species distribution models. For
point estimates the joint distribution models are roughly equivalent to stacked
SDMs, but the JSDMs provide much better estimates of uncertainty [@harris2015;
others?]. In fact, JSDMs and modeling richness directly provide some of the best
estimates of uncertainty across all modeling approaches. This highlights the
importance of evaluating models based on their uncertainty as well as their mean
estimates (Box 1). Not doing so is particularly problematic in this case because the
summed species distribution models are overly confident meaning that using them
would lead to believing that richness would be restricted to a much narrower
range than would actually occur.

For the error metric that evaluates both models' point estimates and their
uncertainty simultaneously, the JSDM and the nonlinear richness model converge
towards the time-series baselines at the longest time-scales assessed. This
convergence at longer time scales highlights the importance of considering how
forecasting method performance changes with the distance into the future being
predicted (Box 1). The value of this approach, and associated forecast horizons,
has been raised for knowing how far into the future a model can be effectively
used for forecasting [@petchy2015], but our results suggest a broader value to
this approach for considering the potential importance of different models and
processes for making forecasts at different scales. [[Should make sure Petchy et
al. don't make this point]] Our results show that for forecasts 1-3 years in the
future, baselines and time-series based approaches outperform all other methods
using integrative metrics. At these time scales, environmental changes are
relatively small and changes in the biota may lag behind changes in the
environment. However, as the timescale of the forecast increases to a decade the
JSDMs and environmental richness models get closer to the accuracy of the
time-series methods and even surpass some of those methods. As the forecast
distance increases the amount of environmental change is expected to increase
and the system will have more time to respond, potentially leading to shifts
towards improved relative performance of the models incorporating environmental
and ecological information. If this shift with forecast distance in the relative
importance of different processes and models continues, this could lead to JSDMs
and environmental richness models outperforming time-series based approaches at
sufficiently long time scales. These results suggest that assessment of how
forecast performance changes with time lag is not only important for determining
how far in the future to forecast, but also for determining models and processes
are most relevant for making forecasts at different time scales.

It is also possible that models that include species' relationships to their
environments or direct environmental constraints on richness may continue to
produce forecasts that are worse than simply assuming the systems are
static. This would be expected to occur if richness in these systems is not
changing over the relevant time scales and therefore simple models of no
directional change are most appropriate. Recent suggestions that local scale
richness in some systems is not changing directionally at multi-decadal scales
supports this possibility [@brown2001; @ernest2001; @dornelas2014;
@vellend2013]. This lack of change may be expected even in the presence of
substantial changes in environmental conditions and species composition at a
site due to the replacement of species from the regional pool [@brown2001;
@ernest2001]. On average the Breeding Bird Survey sites used in this study show
little change in richness ([[Add summary stat here]]) [see also
@lasorte2005]. This absence of rapid change is beneficial for the absolute
accuracy of forecasts across sites, because if a past year's richness is known
it is easy to estimate a future richness, and explains why stacked SDMs perform
relatively well at this task despite failing to capture meaningful
dynamics. However it makes it difficult to improve forecasts relative to simple
baselines, since those baselines are close to representing what is actually
occurring in the system. This suggests that simple time-series models and
baselines should be actively considered for forecasts of richness and other
stable aspects of biodiversity and also suggests that future efforts to
understand and forecast biodiversity should also focus on composition since it
is expected to be more dynamic [@ernest2001; @dornelas2014].

In addition to consideration of the different process models used for
forecasting it is important to consider the observation models. When working
with any ecological dataset there are imperfections in sampling that have the
potential to influence results. With large scale survey and citizen science
datasets like the Breeding Bird Survey these issues are potentially magnified by
the large number of different observers and major differences in habitat and
species [@sauer1994]. We included an observation model for one of the two major
observation issues known for the Breeding Bird Survey, differences among
observers. Accounting for differences in observers reduced the average error in
point estimates and also improved the coverage of the confidence intervals. In
addition, it resulted in changes in which models performed best, most notably
reducing the relative performance of the naive model for point estimates. This
suggests that the naive model performed well in part because it was capable of
accommodating rapid shifts in estimated richness introduced by changes in the
observer. These kinds of rapid changes were difficult for the other time-series
models to accommodate and so the average performance of the ARIMA and average
models improved once this source of observation error was addressed. This
demonstrates that properly modeling observation error can be important for
properly estimating and reducing uncertainty in forecasts and can also lead to
changes in the best methods for forecasting. We did not address differences in
detection probability across species and sites [@boulinier1998] since there is
no clear way to address this issue without making strong assumptions about the
data (i.e., assuming there is no biological variation in stops along a route),
but this would be a valuable addition to future forecasting models.

Future efforts will also need to begin to address the additional uncertainty
that comes from error in forecasting the environmental conditions themselves. In
this and other hindcasting studies the environmental conditions for the "future"
are known because the data has already been observed. In real forecasts the
environmental conditions themselves have to be forecast and those forecasts have
uncertainty and bias. This will cause ecological forecasts that use
environmental data to be more uncertain than observed in current hindcasting
efforts. It is important to correctly incorporate this uncertainty in the
predictor variables into forecasting models [@clark2001; @dietz2017]. Difficulty
in forecasting future environmental conditions at small scales will present
continued challenges for models incorporating these conditions and this may
result in a continued advantage to simple time-series based approaches.

Another challenge for ecological forecasting is the timescales of typical
forecasts. We evaluated forecasts using yearly timesteps up to a decade into the
future. In contrast, many ecological forecasts are temporally aggregated to 5-30
year timesteps and projected up to a century into the future. These are commonly
beyond the career or even lifespan of the researchers, making reasonable model
assessment impossible. Currently assessments at these long time-scales can be
only be made with a small number of opportune datasets. For examples hindcasts
have been used to asses models of species richness [@alger2009, distler2015] and
distribution [@rappacciuolo2012, @moran-ordonez2017, @araujo2005,
@eskildsen2013] using data aggregated across similar time scales. More
evaluation studies like these are needed but there is a paucity of adequate long
term data available. Short term forecasts will allow for relatively rapid
validation of models by researchers without needing to wait for enough data to
accumulate [tredennick2016; @dietze2016]. This does not mean that they should
replace large scale long-term forecasts. As discussed above, drivers of species
richness are likely to differ at different temporal scales [@rosenzweig1995,
@white2004, @white2007]. Thus short-term forecasts should be expected to add
value to end users and inform, not replace, long-term forecasts. Hopefully by
evaluating trends in the error shorter term forecasts as done here it will be
possible to also use the information from short-term forecasting to improve
longer-term estimates.

The science of forecasting biodiversity, and ecology more broadly, remains in
its infancy and it is important to consider the general inability the
forecasting methods to improve on simple baselines in that context. When weather
forecasting first started the forecasts were likewise worse than simple
baselines [@mcgill2012]. One of the things that helped weather forecasts improve
was large numbers of forecasts were made in public, which allowed different
approaches to forecasting to be regularly assessed and improved [@mcgill2012;
@silver2012]. This suggests that it is important for ecologists to start
regularly making and evaluating true ecological forecasts, even if they perform
poorly, and to make these forecasts publicly available for assessment. These
forecasts should include both short-term predictions, which can be assessed
quickly, and mid to long-term forecasts to help assess long time-scale processes
and determine how far into the future we can successfully forecast
[@tredenick2016; @dietz2017]. Forecasts based on the models in this paper from
now until 2050 are openly archived on Dryad so that we and others can assess how
well they perform and we plan to evaluate these forecasts and report the results
as each new year of BBS data becomes available. Weather forecasting has
continually improved throughout its history [@bauer2015] in part due to making
and evaluating public forecasts [@mcgill2012] and we hope this will help ecology
do the same.

Making successful ecological forecasts will be challenging. Ecological systems
are complex, fundamental theory is less refined than simpler physical/chemical
systems, and we currently lack the large amounts of data that produce effective
forecasts through machine learning. Despite this we believe that progress can be
made if we build an active forecasting culture in ecology that builds and
assesses forecasts in ways that will allow us to improve the effectiveness of
ecological forecasts most rapidly (Box 1). This includes expanding the scope of
the ecological and environmental data we work with, paying attention to
uncertainty in both model building and forecast evaluation, and rigorously
assessing forecasts using a combination of hindcasting, archived forecasts, and
comparisons to simple baselines.

@@

Typically ecological forecasts use one modeling approach or a small number of
related approaches. By fitting and evaluating multiple modeling approaches we
can learn more rapidly about the best approaches for making predictions for a
given ecological quantity. This includes comparing process based and data-driven
models and comparing the accuracy of forecasts to simple baselines to determine
if the modeled forecasts are more accurate than the naive assumption that the
world is static [@ye2015].

@@

Forecasts describe how systems are expected to change through time. While some
areas of ecological forecasting focus primarily on time-series data, others
primarily focus on using spatial models and space-for time substitutions. Using
ecological and environmental time-series data allows the consideration of actual
dynamics from both a process and error structure perspective [@treddenick2016a].

@@

Understanding uncertainty in a forecast is just as important as understanding
the average or expected outcome. Failing to account for uncertainty can result
in overconfidence in highly uncertain outcomes leading to poor decision making
and erosion of confidence in ecological forecasts. Models should explicitly
include sources of uncertainty and propagate them through the forecast where
possible [@clark2001; @dietz2017]. Evaluations of forecasts should assess the
accuracy of uncertainties as well as point estimates [@dietz2017].

@@

Many ecological forecasts use data that is readily available and easy to work
with. While ease of use is a reasonable consideration it is also important to
include predictor variables that are expected to relate to the ecological
quantity being forecast and dynamic time-series of predictors instead of
long-term averages. Investing time in identifying and acquiring better predictor
variables may have at least as many benefits as using more sophisticated
modeling techniques. [[Need a citation or two here]]

### 5. Address unknown or unmeasured predictors

Ecological systems are complex and many biotic and abiotic aspects of the
environment are not regularly measured. As a result, some sites may deviate in
consistent ways from model predictions. Unknown or unmeasured predictors can be
incorporated in models using site-level random effects (potentially spatially
autocorrelated) or by using latent variables that can identify unmeasured
gradients [@harris2015].

### 6. Assess how forecast accuracy changes with time-lag

@@

### 7. Include an observation model

Ecological observations are influenced by both the underlying processes and how
the system is sampled. When possible forecasts should model the factors
influencing the observation of the data. [[citation?]]

### 8. Validate using hindcasting

To evaluate the expected accuracy and uncertainty of forecasts assess the
performance of these forecasts within existing time-series data. [[citaton?]]

### 9. Publicly archive forecasts

To allow the future evaluation of the accuracy and uncertainty of forecasts the
forecast values and/or models should be archived so that they can be assessed
after new data is generated [@mcgill2012; @silver2012]. Enough information
should be provided to allow an unambiguous assessment of the forecast
performance.

### 10. Make short-term and long-term predictions

In cases where long-term predictions are the primary goal, short-term should
also be made to accommodate the time-scales of planning and management decisions
and to allow the accuracy of the forecasts to be quickly evaluated
[@treddenick2016].
